---
title: "Simulating Exponentials"
author: "Jade Forlani-Brennan"
date: "10/02/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview
This project report is in 2 sections:

1. Exponential Distrubtion Sampling Simulations
2. Analysing Tooth Growth Data

In the first section of this report I simulate an exponential distrubtion with
parameter lambda of 0.2. And repeat this simulation 40 times, taking the mean
and variance of each of the 40 simulations. The theoretical and sample mean is
compared. And the same for the theoretical and sample variance. Lastly the 
normal distrubtion shape of the samples is shown.

In the second section Tooth Growth Data is explored and summarised. Finally, 
tooth growth by supp and dose is compared with confidence intervals.

# Exponential Distrubtion Sampling Simulations

```{r expConstants}
lambda = 0.2 # number
n = 1000 #  size of each distribution, number of data points
B = 40 # number of simulations
```

## Simulation

First, creating the histogram of a single simulation consisting of **`r n`**
data points.
```{r expSimulation}
hist(rexp(n,rate = lambda))
```

We see the tell tale shape of an expoentially decaying distrubtion.
Most of the values generated are closer to 0, with less and less larger numbers.

Next, repeating this simulation **`r B`** times and looking at the mean for each
simulation and storing ths result in the means variable.

```{r expMeans}
means = NULL
for (i in 1:n) means = c(means, mean(rexp(B,rate=lambda)))
```

This is then repeated again **`r B`** times this time looking at the variance
of each simulation, storing this result in the vars variable.

```{r expVars}
vars = NULL
for (i in 1:n) vars = c(vars, var(rexp(B,rate=lambda)))
```


## Sample Mean vs. Theoretical Mean

```{r expMean}
meanTheory = 1/lambda # theoretical mean
meanSample = means[1] # sample mean
meanAvg = mean(means) # average of sample means
```
The theoretical mean is **`r meanTheory`** vs the sample mean of **`r meanSample`**.
Which is very close with a single sample of **`r n`** data points but not exact.
However, if we average all the samples out the average simulated mean is 
**`r meanAvg`** which is very close to the theoretical number. However, still
slightly difference because of simulation or *"Monte Carlo"* error.

## Sample Variance vs. Theoretical Variance

```{r expVar}
sdTheory = 1/lambda # theoretical stmeanTheoryandard deviation
varTheory = sdTheory^2 # theoretical variance
varSample = vars[1] # sample variance
varAvg = mean(vars) # average of sample variances
```
The theoretical variance is the sample standard deviation squared equal to **`r varTheory`**.
Compare this with the sample variance of **`r varSample`**.
Which is close but not the sample despite a single sample of **`r n`** data points.
However, if we average all the samples out the average simulated variane is 
**`r varAvg`** which is very close to the theoretical number. However, still
slightly difference because of simulation or *"Monte Carlo"* error.

## Sample Means Normally Distrubted

Looking at a histogram of the mean of each of the **`r B`** simulations we see 
what looks much closer to a normal distrubtion shape.

```{r expNormalMeans}
hist(means,
     breaks = 20,
     main = paste("Histogram of", B, "exponential distrubtion simulations with",n,"data points"),
     xlab = "Simulation Mean")
```



We can tell thi from the tell tale bell curve shape centered at approximatly
the theoretical mean of **`r meanTheory`** which thins out at the tails.

# Analysing Tooth Growth Data

## Exploratory Data Analysis

## Summary

## Hypothesis Testing

## Assumptions & Conclusions

```{r }

```
